---
title: "STAT 3011 Week 9"
subtitle: "Section 007"
date: "March/25/2026"
output: html_document
header-includes:
        \usepackage{fancyhdr}
        \pagestyle{fancy}
        \chead{Week 9}
        \rhead{STAT 3011}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Optional: Identifying Missing Values in Data

Lets say you have a vector with some missing data

```{r}
x = c(1, -4.2, NA, 7.2)
is.na(x)
```

```{r}
sum(!is.na(x))
```

Finally, to get the number of data points that is not NA

```{r}
sum(!is.na(x))
```

## How Much Do UMN Students Workout Each Day?

We will keep using the “Getting to Know You” data. Assume that the “Getting To Know You” Survey is a random sample drawn from the population of all U of M students.

Download the csv file of the survey data from Canvas and import it into R.

```{r}
Getting2NoU <- read.csv(
  "https://raw.githubusercontent.com/Talha-Hamza/STAT3011-015-S26/refs/heads/main/Getting2NoU_S2026.csv",
  header = TRUE)
```

Suppose we are interested in providing an interval estimate for $\mu$, the mean amount of time that UMN students work out each day. Let’s construct a 96% confidence interval for $\mu$.

The sample data are stored in the column of Workout (how many minutes per day a student exercises or participants in sports.)

## Checking Conditions for Confidence Interval

1.) **Do we have independent observations?** Yes, since the sample is assumed to be random drawn from a large population.

2.) **Do we have a normal population**

```{r}
hist(Getting2NoU$Workout)
```

```{r}
```

No, the population is likely right skewed.

**Is the sample large enough?**

```{r}
n <- sum(!is.na(Getting2NoU$Workout))
n
```

Since n is greater than 30, yes, the sample size is large enough.

## Constructing the CI “By Hand”

**a.) Point Estimate** $\bar{x}$

```{r}
x_bar <- mean(Getting2NoU$Workout, na.rm = TRUE)
x_bar
```

**b.) Standard Error of** $\bar{x}$ Recall the formula for standard error of $\bar{x}$ is $\dfrac{s}{\sqrt{n}}$

```{r}
s <- sd(Getting2NoU$Workout, na.rm = TRUE)
se <- s/sqrt(n)
se

```

**c.) Critical Value of** $t^*$

Since we want the 96% CI, $\alpha = 0.04$ Recall from the slide $t^*$ = qt(1-$\alpha/2$), df - n-1}

```{r}
alpha = 1-0.96
t_star <- qt(1-alpha/2, df=n-1)
t_star
```

Note: for very large samples (such as ours) the t-distribution is very close to the standard normal distribution N\~(0,1)

```{r}
qt(1-alpha/2, df = 999) # t-distribution's 98th percentile
```

```{r}
qnorm(1-alpha/2) # Standard normal distributions' 98th percentile
```

**d.)Assembling the Components of CI**

Recall from slides the CI formula: $\overline{x} \pm t^* \cdot \underbrace{\dfrac{s}{\sqrt{n}}}_{\text{standard error}}$

```{r}
x_bar +c(-1,1)*t_star*se
```

Interpret: We are 96% confident that the true average time students spend per day exercising is between 54.1 min to 62.8 min.

## Constructing the CI By t.test( )

```{r}

```

## Exercise

**Your turn: construct a 94% CI for the average height of UMN students by hand and by t.test( )?**

```{r}


```

```{r}

```
